---
title: "Final_Project_Code_2025"
output: html_document
date: "2025-05-17"
---
# Load libraries
```{r}
library(tidyverse)
library(forecast)
library(tseries)
library(readxl)
library(MonteCarlo)
library(triangle)
library(reshape2)
```

# Import and read in rainfall data from Bureau of Meteorology (BOM) which has been put into an excel. 
```{r}
rainfall_data <- read_excel("BOMdata.xlsx")
head(rainfall_data)
tail(rainfall_data)
```

# Create a date column and convert to data type. 
```{r}
rainfall_data <- rainfall_data %>%
  mutate(Date = as.Date(paste(Year, Month, "01", sep = "-")))
```

# Convert monthly data to weekly format
```{r}
rainfall_data$Week <- format(rainfall_data$Date, "%Y-%U")
weekly_rainfall <- rainfall_data %>%
  group_by(Week) %>%
  summarise(Weekly_mm = sum(Rainfall_mm, na.rm = TRUE))
```

# Convert weekly data to a time series format and remove NA
```{r}
rainfall_weekly_ts <- ts(weekly_rainfall$Weekly_mm, frequency = 52)
rainfall_weekly_ts <- na.omit(rainfall_weekly_ts)
```

# Check stationarity using Augmented Dickey-Fuller test
```{r}
adf_result <- adf.test(rainfall_weekly_ts)
print(adf_result)
#A transformation was not needed as the p value < 0.05. 
```

# Fit the SARIMA model
```{r}
auto_model <- auto.arima(rainfall_weekly_ts, seasonal = TRUE)
summary(auto_model)
```

# Forecast the next 52 weeks (1 year) and plot
```{r}
forecast_rain <- forecast(auto_model, h = 52)
print(forecast_rain)
autoplot(forecast_rain) +
  labs(title = "52-Week Rainfall Forecast", y = "Rainfall (mm)")
```

## Water balance
# Define constants
```{r}
roof_area <- 200
runoff_efficiency <- 0.85
household_size <- 4
greywater_per_capita <- 100
daily_demand_per_person <- 164
```

# Weekly greywater and demand
```{r}
greywater_data <- rep(household_size * greywater_per_capita * 7, length.out = 52)
daily_demand <- household_size * daily_demand_per_person
weekly_demand <- rep(daily_demand * 7, length.out = 52)
```

# Generate a seasonal pattern for water demand
```{r}
seasonal_pattern <- 1 + 0.2 * sin(2 * pi * (1:52) / 52)
adjusted_weekly_demand <- weekly_demand * seasonal_pattern
```

# Convert rainfall to rainwater volume
```{r}
rainwater_volume <- function(rainfall_mm, runoff_eff = runoff_efficiency) {
  return(rainfall_mm * roof_area * runoff_eff)
}
```

# Water balance simulation function
```{r}
simulate_tank <- function(tank_size, rainfall_data, greywater_data, demand_data, runoff_eff = runoff_efficiency) {
  tank_storage <- numeric(length = length(rainfall_data))
  mains_water_used <- numeric(length = length(rainfall_data))
  overflow <- 0
  shortage <- 0
  
  for (week in 1:length(rainfall_data)) {
    rainwater_vol <- rainwater_volume(rainfall_data[week])
    total_input <- rainwater_vol + greywater_data[week]
    prev_storage <- ifelse(week == 1, 0, tank_storage[week - 1])
    tank_storage[week] <- min(tank_size, prev_storage + total_input)
    
    if (tank_storage[week] < demand_data[week]) {
      mains_water_used[week] <- demand_data[week] - tank_storage[week]
      tank_storage[week] <- 0
    } else {
      tank_storage[week] <- tank_storage[week] - demand_data[week]
    }
    
    if (tank_storage[week] == tank_size) {
      overflow <- overflow + (total_input - (tank_size - prev_storage))
    }
  }
  
  return(list(storage = tank_storage, mains_water_used = mains_water_used, overflow = overflow, shortage = shortage))
}
```

# Define tank sizes (litres)
```{r}
tank_sizes <- c(2000, 5000, 10000, 3000, 7000, 8000, 4000)
results <- lapply(tank_sizes, function(size) simulate_tank(size, forecast_rain$mean, greywater_data, adjusted_weekly_demand))
```

# Create dataframes for each tank size and combined data
```{r}
create_result_df <- function(result_list, size) {
  data.frame(
    Week = 1:52,
    TankSize = size,
    Storage = result_list$storage,
    Mains = result_list$mains_water_used,
    Overflow = result_list$overflow
  )
}

df2000 <- create_result_df(results[[1]], 2000)
df5000 <- create_result_df(results[[2]], 5000)
df10000 <- create_result_df(results[[3]], 10000)
df3000 <- create_result_df(results[[4]], 3000)
df7000 <- create_result_df(results[[5]], 7000)
df8000 <- create_result_df(results[[6]], 8000)
df4000 <- create_result_df(results[[7]], 4000)

combined_results <- bind_rows(df2000, df5000, df10000, df3000, df7000, df8000)

```

# Plot combined data
```{r}
combined_long <- combined_results %>%
  select(Week, TankSize, Storage, Mains, Overflow) %>%
  pivot_longer(cols = c(Storage, Mains, Overflow), 
               names_to = "Parameter", 
               values_to = "Value")

#Plot mains, overflow and storage in litres over 52 weeks by tank size
ggplot(combined_long, aes(x = Week, y = Value, color = factor(TankSize))) +
  geom_line(linewidth = 1) +
  facet_wrap(~ Parameter, scales = "free_y") +
  labs(title = "Water System Parameters over 52 Weeks by Tank Size",
       x = "Week",
       y = "Value (Litres)",
       color = "Tank Size (L)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Cost Analysis and Bayesian Monte Carlo Simulation
# Cost parameters
```{r}
installation_costs <- list("2000" = 400, "5000" = 800, "10000" = 1500, "3000" = 550, "7000" = 1000, "8000" = 1200, "4000" = 650)
mains_water_cost_per_kL <- 2
discount_rate <- 0.05
years <- 10
weeks_per_year <- 52
```

# Bayesian Monte Carlo simulation 
```{r}
monte_carlo_simulation <- function(tank_size, n = 1000) {
  results <- data.frame(TankSize = rep(tank_size, n), NPV = numeric(n), CostPerLitre = numeric(n), Payback = numeric(n), PaidBack = logical(n))
  
  for (i in 1:n) {
    greywater_gen <- rnorm(1, mean = 100, sd = 10)
    weekly_greywater <- rep(household_size * greywater_gen * 7, 52)
    demand <- rtriangle(52, a = 100 * 7 * household_size, b = 180 * 7 * household_size, c = 164 * 7 * household_size)
    runoff_eff <- rbeta(1, shape1 = 5, shape2 = 1.2)
    tank_cost_per_litre <- runif(1, 0.5, 1.0)
    maintenance_cost <- rnorm(1, mean = 100, sd = 20)
    
    sim_result <- simulate_tank(tank_size, forecast_rain$mean, weekly_greywater, demand, runoff_eff)
    
    total_mains_saved <- sum(sim_result$mains_water_used)  
    cost_mains_saved <- total_mains_saved * mains_water_cost_per_kL
    capital_cost <- tank_cost_per_litre * tank_size + installation_costs[[as.character(tank_size)]]
    maintenance_total <- maintenance_cost * years
    
    npv_savings <- sum((cost_mains_saved / (1 + discount_rate) ^ (1:years)))
    total_cost <- capital_cost + maintenance_total
    
    if (total_mains_saved > 0) {
      results$NPV[i] <- npv_savings - total_cost
      results$CostPerLitre[i] <- total_cost / total_mains_saved
      results$Payback[i] <- total_cost / (cost_mains_saved / years)
      results$PaidBack[i] <- results$NPV[i] > 0
    } else {
      results$NPV[i] <- NA
      results$CostPerLitre[i] <- NA
      results$Payback[i] <- NA
      results$PaidBack[i] <- FALSE
    }
  }
  
  return(results)
}

```

# Run Monte Carlo simulation for all tank sizes
```{r}
monte_results <- bind_rows(lapply(tank_sizes, function(size) monte_carlo_simulation(size, n = 1000)))

summary_means <- monte_results %>%
  filter(!is.na(NPV)) %>%
  group_by(TankSize) %>%
  summarise(
    Mean_NPV = mean(NPV),
    Mean_CostPerLitre = mean(CostPerLitre),
    Mean_Payback = mean(Payback),
    Proportion_PaidBack = mean(PaidBack)
  )

print(summary_means)
```

# Visualise cost metrics
```{r}
#Plot mean NPV by tank size
ggplot(summary_means, aes(x = factor(TankSize), y = Mean_NPV)) +
  geom_col(fill = "steelblue") +
  labs(title = "Mean Net Present Value by Tank Size",
       x = "Tank Size (L)", y = "Mean NPV ($)") +
  theme_minimal()

#Plot mean cost per litre by tank size
ggplot(summary_means, aes(x = factor(TankSize), y = Mean_CostPerLitre)) +
  geom_col(fill = "darkorange") +
  labs(title = "Mean Cost per Litre of Water Saved",
       x = "Tank Size (L)", y = "Cost per Litre ($/L)") +
  theme_minimal()

#Plot mean payback period by tank size
ggplot(summary_means, aes(x = factor(TankSize), y = Mean_Payback)) +
  geom_col(fill = "forestgreen") +
  labs(title = "Mean Payback Period by Tank Size",
       x = "Tank Size (L)", y = "Years") +
  theme_minimal()

#Plot proportion of simulations where the system was paid back 
ggplot(summary_means, aes(x = factor(TankSize), y = Proportion_PaidBack)) +
  geom_col(fill = "purple") +
  labs(title = "Proportion of Simulations Where System Paid Back",
       x = "Tank Size (L)", y = "Proportion Paid Back") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

```

